# Analysis
## Plan
- Load standardized means
- Fit model using weights for counts inversely related to mean value 
	- Poisson Like: W = 1/(mean +1) since $E(X) = \lambda = \var(X)$
# Install packages
```{r}
if (!requireNamespace("readr", quietly = TRUE))
    install.packages("readr")

## eivreg
if (!requireNamespace("eivtools", quietly = TRUE))
    install.packages("eivtools")

if (!requireNamespace("Biostrings", quietly = TRUE))
    BiocManager::install("Biostrings", lib="/home/mikeg/R/x86_64-pc-linux-gnu-library/3.6")



```

# Load Libraries
```{r}
library(Biostrings) ## process first to avoid conflicts
library(tidyr)
library(tibble)
library(readr)
library(dplyr)
library(stringr)
library(eivtools)

```

# Load and Shape Data

## E-MTAB
```{r}
flatData  <- readr::read_tsv("E-MTAB-2812-query-results.tpms.tsv", skip=4) %>%
    dplyr::rename(ID = `Gene ID`, gene=`Gene Name` )

## use pivot_longer command (not gather which is depricated)
tData <- flatData %>% pivot_longer(-c(ID, gene), names_to = "long_description", values_to = "count")

## Now separate entry in descriptor column into separate column entries and filter out tissue specific data
sData  <- separate(tData, long_description, into=c("sex", "tissue", "stage"), sep=", ") %>% filter(tissue=="organism") %>%
    mutate_if(is.character,as.factor)  %>% # convert char columns to factors
    replace_na(list(count = 0)) # replace NAs with 0.
                                        #Before doing so the min value in the tibble was 0.1


# Create a new set of measurements for 'hermaphrodite embryo Ce' stage
## Get hermphrodite embryo stage names (male, 'embryo Ce' lacks a space at the start)
embryoStageNames  <- levels(sData$stage)[grepl(' embryo Ce', levels(sData$stage) )]
embryoStageCounts  <-  filter(sData, stage %in% embryoStageNames) %>% group_by(ID)

combinedEmbryoStageCounts <- unique(embryoStageCounts %>% summarise(gene, sex, tissue, stage=as.factor("embryo Ce"), median=median(count)))

nonEmbryoStageCounts <- filter(sData, !(stage %in% embryoStageNames)) 

stageCounts <- bind_rows(combinedEmbryoStageCounts, nonEmbryoStageCounts)
## 
## Make things simple and focus on just the 'main' lifestage measurements
## note that the stage "embryo Ce" is specific to sex = male
## Thus need to combine embryo measurements some how
## Try using median of stages

mainLifeStages <- c("embryo Ce", "L1 larvae Ce", "L2 larvae Ce", "L3 larvae Ce", "L4 larvae Ce", "adult Ce", "dauer larvae Ce", "post dauer stage Ce")

mainLSData <- filter(sData, (stage %in% mainLifeStages & sex == "hermaphrodite")) 

summary(mainLSData$count)

```

## ROC 

### Get gene names

#### BROKEN - Don't use Lu's file
```{r}

## NOTE: phiNames is corrupt.  It has replaced some gene names with date formats (e.g. apr-1 has been converted to 1-Apr
phiNamesBroken <- read_csv("phi.mean.by.names.csv", col_names = c("gene","phi2") )
phiNamesBroken[phiNamesBroken$gene=='1-Apr',]

```

#### Extract from FASTA file
```{r}
## Load original FASTA file
## Using Biostrings function which is not a standard df
seqData <- readDNAStringSet("c_elegan.fasta")

## names are really long descriptions.
## NEed to extract relevant part
seqDesc  <- names(seqData)

rawStrExtract <- str_extract(seqDesc, "\\[gene=[A-z0-9\\-.]*\\]")

## verify there's a match for each entry
if(sum(is.na(rawStrExtract)) ==0) print("Every entry matches")

## Verify that all 'names' are unique.
length(seqDesc)
length(unique(seqDesc))

## It appears that
```

#### Get estimates from ROC
```{r}
## detailed information on phi: posterior mean, posterior mean of log10(phi), etc
## StdError really StdDev of posterior
phiPosteriorInfo <- read_csv("phi.posterior.unlabeled.csv") %>% rename(phi = PHI, logPhi = log10.PHI, sd = Std.Error, logSd = log10.Std.Error)


phiDataCheck  <- bind_cols(phiNames, phiPosteriorInfo)
## Verify that phi values line up between the two datasets
plot(phiDataCheck$phi2, phiDataCheck$phi)

## Get rid of phi2 column
phiData <- select(phiDataCheck, -phi2) %>%
    mutate_if(is.character,as.factor)  # convert char columns to factors


## Look at structure of data

## Check for duplicate gene names
phiData[phiDataCheck$name=="wwp-1", ]

## There are 4 entries in the "phi.mean.by.names.csv" with the name "wwp-1" 
## This is because the FASTA file contains 4 isoforms of this gene
## Problem: there's only one WB Gene ID and, thus, only one measurement.
## Solution: Combine separate estimates using median values of phi

combinedPhiData  <- phiData %>% group_by(gene) %>% summarize(median_phi = median(phi), sd_phi = median(sd), isoforms = length(phi) )



```


# Modeling error in RNA Seq Data
If the underlying distribution is 
	- poisson, then Var=Rate so W = 1/count or 1/(count + k)
	
# Fit Using *eivtools* 
```{r}

cData  <-bind_cols(  
eivreg(

```
