---
title: "Plot Output from Fit ROC to Sequence Data"
author: "Michael A. Gilchrist"
date: "28 Jul 2020"
output: pdf_document
---

	
# Background
- This file and `../28_Plot.Output.from.Fit.ROC.to.Sequence/plot.output.from.roc.Rmd` were created from the `updateROC.R` file that Lu has been using.
- Shifting from using filtered fasta files to filtering genome object after initializing it.

# ESS Commands
See (~/Software/R/r.notes.Rmd)  for more details

# Purpose
- Fit ROC to data using easier to understand and modify code
- Save output for future plotting.

# Load Libraries
```{r}

## Clear out any previously defined objects and parameters
rm(list=ls())

## Load Needed Libraries
library(tidyr)
library(tibble)
library(readr)
library(dplyr)
library(stringr)
library(forcats)
library(knitr)
library(purrr)
library(AnaCoDa, verbose = TRUE, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6") ## Could use Version(); if it's not there build locally using script
## library(profmem, verbose = TRUE)  #is this still necessary?
library(ggplot2)
library(ggpubr)
library(ggpmisc)
library(gridExtra)
library(viridis)
library(latex2exp)
      ## Usage: TeX(sprintf("$\\alpha = %d$", alpha)
      ##        TeX('$\\alpha  x^\\alpha$, where $\\alpha \\in 1\\ldots 5$')

```

# Define Local Functions
```{r}
## THis was written by Alex Cope
createTracePlots <- function(trace, model, genome, numMixtures, samples, mixture.labels, samples.percent.keep=1)
{
  for (i in 1:numMixtures)
  {
    plot(trace, what = "Mutation", mixture = i)
    plot(trace, what = "Selection", mixture = i)
    plot(model, genome, samples = samples*samples.percent.keep, mixture = i,main = mixture.labels[i])
  }
}

```

# Load Output From ROC

```{r}


## Information on genes based on FASTA file
load("./Input/all.nuclear.and.monoform.gene.lists.Rdata")
rocOutputDir <- "./Output-28_Fit.ROC.to.Sequences/"

## Wrong way to load objects, but not they do provide some information
tmp <- load("Output-28_Fit.ROC.to.Sequences/R.Objects/parameter.round-4.Rda")
tmp

## Load settings for ROC fit
load(paste0(rocOutputDir, "parameterSaveList.Rdata"))

## Load objects parameter.round-X.Rda

data.dir <- paste0(rocOutputDir, "R.Objects/")

#fs::dir_ls(regexp = "parameter.round-[0-9]+.Rda", path = data.dir)

parameterSet <-
    sapply(1:max.round, FUN = function(x){
    loadParameterObject(paste0(data.dir, "parameter.round-", x, ".Rda")
         )
    
    }
    )

mcmcSet <-
    sapply(1:max.round, FUN = function(x){
    loadMCMCObject(paste0(data.dir, "mcmc.round-", x, ".Rda")
         )
    
    }
    )

## Functions available for 'parameter' Object
methods(class = "Rcpp_ROCParameter")
methods(class = "Rcpp_ROCModel")
methods(class = "Rcpp_ROCTrace")
```


## Plotting Routines

```{r}


parameter <- parameterSet[[10]]
## This works!
## what=c("Mutation", "Selection", "MixtureProbability" ,"Sphi", "Mphi", "Aphi", "Sepsilon", "ExpectedPhi", "Expression","NSEProb","NSERate","InitiationCost","PartitionFunction")
trace <- parameter$getTraceObject()
plot(trace, what = "ExpectedPhi")
plot(trace, what = "Sphi")
plot(trace, what = "Mutation")
plot(trace, what = "Selection")
plot(trace, what = "Expression")


## Plot ACF
numMixtures = 1

par(mfrow=c(5,4), new=TRUE) 
acfCSP(parameter,
       csp="Selection",
       numMixtures = numMixtures,
       samples=samples.to.keep)


par(mfrow=c(5,4), new=TRUE) 
acfCSP(parameter,
       csp="Mutation",
       numMixtures = numMixtures,
       samples=samples.to.keep)


## Plot Acceptance Rates
aa <- aminoAcids()

aa <- aa[!(aa %in% c("M", "X", "W"))]
trace$getCodonSpecificAcceptanceRateTraceForAA("A");

tmp  <- lapply(aa, function(x){
    yvals <- trace$getCodonSpecificAcceptanceRateTraceForAA(x)
    plt <- ggplot(y = yvals,
          x = 1:length(yvals) ) + geom_line()
    return(plt)
}
)
grid.arrange(grobs = tmp, ncol=4)


yvals <- trace$getCodonSpecificAcceptanceRateTraceForAA("A")
xvals  <-  1:length(yvals)


tmp  <- lapply(aa, function(x){
    yvals <- trace$getCodonSpecificAcceptanceRateTraceForAA(x)
    xvals  <-  1:length(yvals)
    data <- as_tibble(cbind(window = xvals, acceptance.rate = yvals))
    plt <- qplot(data = data,
                 y = acceptance.rate,
                 x = window,
                 geom = "line",
                 ylim = c(0, 0.6),
                 main = x) +
        geom_hline(yintercept = 0.2, linetype = 2) +
        geom_hline(yintercept = 0.35, linetype = 2) 
    return(plt)
}
)

grid.arrange(grobs = tmp, ncol=4)


## Plot Model
## Does this create codon freq vs. log(phi) plots?
## If so, where does it get the information from?
## I thought the samples were stored in the parameter object
plot(model, genome,
     samples = samples*samples.percent.keep,
     mixture = i,
     main = mixture.labels[i])

plot(mcmcSet[[1]], what = "LogPosterior")



for(a in aa)
{
  if (a=="M"||a=="X"||a=="W") next
  accept.trace <- trace$getCodonSpecificAcceptanceRateTraceForAA(a)
  len <- length(accept.trace)
  mean.acceptance <- mean(accept.trace[(len-len*0.5):len])
  if (mean.acceptance < 0.1 || mean.acceptance > 0.44) done.adapt <- FALSE
  plot(accept.trace,main=paste0("Acceptace Rate for ",a), xlab="Samples",ylab="Acceptance Rate",type="l")
}



## Export plots
pdf(paste0(output.dir,"Graphs/CSP_traces_CUB_plot.pdf"), width = 11, height = 12)

## This is a locally defined function.
createTracePlots(trace=trace,model=model,genome=genome,numMixtures=numMixtures,samples=samples,samples.percent.keep = 1,mixture.labels = mixture.labels)

dev.off()


## pdf(paste(dir_name,"Graphs/Parameter_comparisons.pdf",sep="/"), width = 11, height = 12)

## This fails with "Error in x$getCodonSpecificPosteriorMean(mixture, samples, codons[i],  :  could not find valid method"
## I believe it's set up to compare two mixtures and since we only have one it fails.
## How do we do a plot this to compare two parameter objects?
plot(parameter, what="Mutation",
     mixture.name = "round",
     samples=samples.to.keep)

plot(parameter,what="Selection",
     samples=samples*percent.to.keep,
     mixture.name=mixture.labels)
dev.off()

trace <- parameter$getTraceObject()


#plots different aspects of trace

pdf(paste(dir_name,"Graphs/mcmc_traces.pdf",sep="/"))
plot(mcmc,what = "LogPosterior")
plot(trace, what = "ExpectedPhi")
aa <- aminoAcids()
done.adapt <- TRUE
for(a in aa)
{
  if (a=="M"||a=="X"||a=="W") next
  accept.trace <- trace$getCodonSpecificAcceptanceRateTraceForAA(a)
  len <- length(accept.trace)
  mean.acceptance <- mean(accept.trace[(len-len*0.5):len])
  if (mean.acceptance < 0.1 || mean.acceptance > 0.44) done.adapt <- FALSE
  plot(accept.trace,main=paste0("Acceptace Rate for ",a),xlab="Samples",ylab="Acceptance Rate",type="l")
}
acfCSP(parameter,csp="Selection",numMixtures = numMixtures,samples=samples*percent.to.keep)
acfCSP(parameter,csp="Mutation",numMixtures = numMixtures,samples=samples*percent.to.keep)
dev.off()


for (i in 1:numMixtures)
{
  param.diag<-convergence.test(trace,samples=samples*percent.to.keep,thin = thinning,what="Selection",mixture=i,frac1=0.1)
  z.scores <- param.diag$z[which(abs(param.diag$z) > 1.96)]
  if (length(z.scores) > 0)
  {
    param.conv <- FALSE
  }
  write(param.diag$z,paste0(dir_name,"/Parameter_est/convergence_delta_eta_",i,".txt"),ncolumns = 1)
}


for (i in 1:numMixtures)
{
  param.diag<-convergence.test(trace,samples=samples*percent.to.keep,thin = thinning,what="Mutation",mixture=i,frac1=0.1)
  z.scores <- param.diag$z[which(abs(param.diag$z) > 1.96)]
  if (length(z.scores) > 0)
  {
    param.conv <- FALSE
  }
  write(param.diag$z,paste0(dir_name,"/Parameter_est/convergence_delta_m_",i,".txt"),ncolumns = 1)
}


writeParameterObject(parameter,paste(dir_name,"R_objects/parameter.Rda",sep="/"))
writeMCMCObject(mcmc,file=paste(dir_name,"R_objects/mcmc.Rda",sep="/"))


diag <- convergence.test(mcmc,samples = samples*percent.to.keep,thin=thinning,frac1=0.2)
z<-abs(diag$z)
done <- (z > 1.96) && param.conv
rm(parameter)
rm(trace)
rm(model)
while((!done) && (run_number <= 3))
{
  parameter<-initializeParameterObject(init.with.restart.file = paste(dir_name,"Restart_files/rstartFile.rst_final",sep="/"),model="ROC")
  run_number <- run_number + 1
  dir_name <- paste0(directory,"/run_",run_number)
  dir.create(dir_name)
  dir.create(paste(dir_name,"Graphs",sep="/"))
  dir.create(paste(dir_name,"Restart_files",sep="/"))
  dir.create(paste(dir_name,"Parameter_est",sep="/"))
  dir.create(paste(dir_name,"R_objects",sep="/"))
  
  mcmc <- initializeMCMCObject(samples=samples, thinning=thinning, adaptive.width=adaptiveWidth,
                               est.expression=T, est.csp=TRUE, est.hyper=F,est.mix=FALSE)
  
  model <- initializeModelObject(parameter, "ROC", with.phi)
  setRestartSettings(mcmc, paste(dir_name,"Restart_files/rstartFile.rst",sep="/"), adaptiveWidth, F)
  sys.runtime <- system.time(
    runMCMC(mcmc, genome, model, num_threads,div=0)
  )
  sys.runtime <- data.frame(Value=names(sys.runtime),Time=as.vector(sys.runtime))
  write.table(sys.runtime,file=paste(dir_name,"mcmc_runtime.csv",sep="/"),sep=",",col.names = T,row.names = T,quote=F)
  
  createParameterOutput(parameter = parameter,numMixtures = numMixtures,samples = samples,mixture.labels = mixture.labels,samples.percent.keep = percent.to.keep,relative.to.optimal.codon = F,report.original.ref = T)
  
  expressionValues <- getExpressionEstimates(parameter,c(1:size),samples*percent.to.keep)
  write.table(expressionValues,file=paste(dir_name,"Parameter_est/gene_expression.txt",sep="/"),sep=",",col.names = T,quote = F,row.names = F)
  
  #
  #plots different aspects of trace
  trace <- parameter$getTraceObject()
  pdf(paste(dir_name,"Graphs/mcmc_traces.pdf",sep="/"))
  plot(mcmc,what = "LogPosterior")
  plot(trace, what = "ExpectedPhi")
  aa <- aminoAcids()
  done.adapt <- TRUE
  for(a in aa)
  {
    if (a=="M"||a=="X"||a=="W") next
    accept.trace <- trace$getCodonSpecificAcceptanceRateTraceForAA(a)
    len <- length(accept.trace)
    mean.acceptance <- mean(accept.trace[(len-len*0.5):len])
    if (mean.acceptance < 0.1 || mean.acceptance > 0.44) done.adapt <- FALSE
    plot(accept.trace,main=paste0("Acceptace Rate for ",a),xlab="Samples",ylab="Acceptance Rate",type="l")
  }
  acfCSP(parameter,csp="Selection",numMixtures = numMixtures,samples=samples*percent.to.keep)
  acfCSP(parameter,csp="Mutation",numMixtures = numMixtures,samples=samples*percent.to.keep)
  dev.off()
  
  
  for (i in 1:numMixtures)
  {
    param.diag<-convergence.test(trace,samples=samples*percent.to.keep,thin = thinning,what="Selection",mixture=i,frac1=0.1)
    z.scores <- param.diag$z[which(abs(param.diag$z) > 1.96)]
    if (length(z.scores) > 0)
    {
      param.conv <- FALSE
    }
    write(param.diag$z,paste0(dir_name,"/Parameter_est/convergence_delta_eta_",i,".txt"),ncolumns = 1)
  }
  
  
  for (i in 1:numMixtures)
  {
    param.diag<-convergence.test(trace,samples=samples*percent.to.keep,thin = thinning,what="Mutation",mixture=i,frac1=0.1)
    z.scores <- param.diag$z[which(abs(param.diag$z) > 1.96)]
    if (length(z.scores) > 0)
    {
      param.conv <- FALSE
    }
    write(param.diag$z,paste0(dir_name,"/Parameter_est/convergence_delta_m_",i,".txt"),ncolumns = 1)
  }
  pdf(paste(dir_name,"Graphs/Parameter_comparisons.pdf",sep="/"), width = 11, height = 12)
  plot(parameter,what="Mutation",samples=samples*percent.to.keep,mixture.name=mixture.labels)
  plot(parameter,what="Selection",samples=samples*percent.to.keep,mixture.name=mixture.labels)
  dev.off()
  
  
  pdf(paste(dir_name,"Graphs/CSP_traces_CUB_plot.pdf",sep="/"), width = 11, height = 12)
  createTracePlots(trace=trace,model=model,genome=genome,numMixtures=numMixtures,samples=samples,samples.percent.keep = percent.to.keep,mixture.labels = mixture.labels)
  dev.off()
  writeParameterObject(parameter,paste(dir_name,"R_objects/parameter.Rda",sep="/"))
  writeMCMCObject(mcmc,file=paste(dir_name,"R_objects/mcmc.Rda",sep="/"))
  
  diag <- convergence.test(mcmc,samples = samples*percent.to.keep,thin=thinning,frac1=0.1)
  z<-abs(diag$z)
  done <- (z > 1.96) && param.conv
  rm(parameter)
  rm(trace)
  rm(model)
}

samples <- 10000
thinning <- 5
parameter<-initializeParameterObject(init.with.restart.file = paste(dir_name,"Restart_files/rstartFile.rst_final",sep="/"),model="ROC")
run_number <- run_number + 1
dir_name <- paste0(directory,"/final_run")
dir.create(dir_name)
dir.create(paste(dir_name,"Graphs",sep="/"))
dir.create(paste(dir_name,"Restart_files",sep="/"))
dir.create(paste(dir_name,"Parameter_est",sep="/"))
dir.create(paste(dir_name,"R_objects",sep="/"))

mcmc <- initializeMCMCObject(samples=samples, thinning=thinning, adaptive.width=adaptiveWidth,
                             est.expression=TRUE, est.csp=TRUE, est.hyper=F,est.mix=FALSE)

mcmc$setStepsToAdapt(0)

model <- initializeModelObject(parameter, "ROC", with.phi)
setRestartSettings(mcmc, paste(dir_name,"Restart_files/rstartFile.rst",sep="/"), adaptiveWidth, F)
#run mcmc on genome with parameter using model
#p<-profmem({
sys.runtime <- system.time(
  runMCMC(mcmc, genome, model, num_threads)
)
sys.runtime <- data.frame(Value=names(sys.runtime),Time=as.vector(sys.runtime))
write.table(sys.runtime,file=paste(dir_name,"mcmc_runtime.csv",sep="/"),sep=",",col.names = T,row.names = T,quote=F)


createParameterOutput(parameter = parameter,numMixtures = numMixtures,samples = samples,mixture.labels = mixture.labels,samples.percent.keep = 1,relative.to.optimal.codon = F,report.original.ref = T)

# mixtureAssignment <- getMixtureAssignmentEstimate(parameter,c(1:size),samples*0.5)
expressionValues <- getExpressionEstimates(parameter,c(1:size),samples)

write.table(expressionValues,file=paste(dir_name,"Parameter_est/gene_expression.txt",sep="/"),sep=",",col.names = T,quote = F,row.names = F)
writeParameterObject(parameter,paste(dir_name,"R_objects/parameter.Rda",sep="/"))
writeMCMCObject(mcmc,file=paste(dir_name,"R_objects/mcmc.Rda",sep="/"))


#plots different aspects of trace
trace <- parameter$getTraceObject()
pdf(paste(dir_name,"Graphs/mcmc_traces.pdf",sep="/"))
plot(mcmc,what = "LogPosterior")
plot(trace, what = "ExpectedPhi")
acfCSP(parameter,csp="Selection",numMixtures = numMixtures,samples=samples)
acfCSP(parameter,csp="Mutation",numMixtures = numMixtures,samples=samples)
dev.off()

pdf(paste(dir_name,"Graphs/Parameter_comparisons.pdf",sep="/"), width = 11, height = 12)
plot(parameter,what="Mutation",samples=samples,mixture.name=mixture.labels)
plot(parameter,what="Selection",samples=samples,mixture.name=mixture.labels)
dev.off()


pdf(paste(dir_name,"Graphs/CSP_traces_CUB_plot.pdf",sep="/"), width = 11, height = 12)
createTracePlots(trace=trace,model=model,genome=genome,numMixtures=numMixtures,samples=samples,samples.percent.keep = 1,mixture.labels = mixture.labels)
dev.off()

for (i in 1:numMixtures)
{
  param.diag<-convergence.test(trace,samples=samples,thin = thinning,what="Selection",mixture=i,frac1=0.1)
  z.scores <- param.diag$z[which(abs(param.diag$z) > 1.96)]
  if (length(z.scores) > 0)
  {
    param.conv <- FALSE
  }
  write(param.diag$z,paste0(dir_name,"/Parameter_est/convergence_delta_eta_",i,".txt"),ncolumns = 1)
}

for (i in 1:numMixtures)
{
  param.diag<-convergence.test(trace,samples=samples,thin = thinning,what="Mutation",mixture=i,frac1=0.1)
  z.scores <- param.diag$z[which(abs(param.diag$z) > 1.96)]
  if (length(z.scores) > 0)
  {
    param.conv <- FALSE
  }
  write(param.diag$z,paste0(dir_name,"/Parameter_est/convergence_delta_m_",i,".txt"),ncolumns = 1)
}
rm(trace)
rm(model)
writeParameterObject(parameter,paste(dir_name,"R_objects/parameter.Rda",sep="/"))
writeMCMCObject(mcmc,file=paste(dir_name,"R_objects/mcmc.Rda",sep="/"))


