# Analysis
## Plan
- Load standardized means
- Fit model using weights for counts inversely related to mean value 
	- Poisson Like: W = 1/(mean +1) since $E(X) = \lambda = \var(X)$
# Install packages
```{r}
if (!requireNamespace("readr", quietly = TRUE))
    install.packages("readr")

if (!requireNamespace("forcats", quietly = TRUE))
    install.packages("forcats")

## eivreg
if (!requireNamespace("eivtools", quietly = TRUE))
    install.packages("eivtools")

if (!requireNamespace("Biostrings", quietly = TRUE))
    BiocManager::install("Biostrings", lib="/home/mikeg/R/x86_64-pc-linux-gnu-library/3.6")



```

# Load Libraries
```{r}
library(Biostrings) ## process first to avoid conflicts
library(tidyr)
library(tibble)
library(readr)
library(dplyr)
library(stringr)
library(forcats)


```

# Load and Shape Data

## E-MTAB
```{r}
emtabFile <- "E-MTAB-2812-query-results.tpms.tsv"
flatData  <- readr::read_tsv(emtabFile, skip=4) %>%
    dplyr::rename(ID = `Gene ID`, gene=`Gene Name` )

## use pivot_longer command (not gather which is depricated)
tData <- flatData %>% pivot_longer(-c(ID, gene), names_to = "long_description", values_to = "count")

## Now separate entry in descriptor column into separate column entries and filter out tissue specific data
sData  <- separate(tData, long_description, into=c("sex", "tissue", "stage"), sep=", ") %>% filter(tissue=="organism") %>%
#    mutate_if(is.character,as.factor)  %>% # convert char columns to factors
    replace_na(list(count = 0)) # replace NAs with 0.
                                        #Before doing so the min value in the tibble was 0.1

creationInfo <- paste0("\tDate: ", date(), "\n\tLocation: ", sub(".*/AcrossTissue", "AcrossTissue", getwd()))
comment(sData) <- paste0("Data from ", emtabFile, "in tidy format.\n", creationInfo) 

# Create a new set of measurements for 'hermaphrodite embryo Ce' stage
## Get hermphrodite embryo stage names (male, 'embryo Ce' lacks a space at the start)
embryoStageNames  <- levels(sData$stage)[grepl(' embryo Ce', levels(sData$stage) )]
embryoStageCounts  <-  filter(sData, stage %in% embryoStageNames) %>% group_by(ID)

combinedEmbryoStageCounts <- unique(embryoStageCounts %>% summarise(gene, sex, tissue, stage="embryo Ce", median=median(count)))

nonEmbryoStageCounts <- filter(sData, !(stage %in% embryoStageNames)) 

stageCounts <- bind_rows(combinedEmbryoStageCounts, nonEmbryoStageCounts)
## 
## Make things simple and focus on just the 'main' lifestage measurements
## note that the stage "embryo Ce" is specific to sex = male
## Thus need to combine embryo measurements some how
## Try using median of stages

mainLifeStages <- c("embryo Ce", "L1 larvae Ce", "L2 larvae Ce", "L3 larvae Ce", "L4 larvae Ce", "adult Ce", "dauer larvae Ce", "post dauer stage Ce")
comment(mainLifeStages)=paste0("List of main lifestages to use in regression.\n", creationInfo)

tidyFullLifeStageData <- filter(sData, (stage %in% mainLifeStages)) ## & sex == "hermaphrodite")) 
comment(tidyFullLifeStageData) <- c("RNASeq counts for each of the\n\t- hermaphrodite non-embryo lifestages: L1 larvae Ce, L2 larvae Ce, ... adult\n\t- male lifestages\n\t- median of the very specific 'hermphrodite,  *embryo Ce' which are 'hermphrodite, embryo Ce.\nAll data for tissue == organism.\nCounts from E-MTAB-2812-query-results.tpms.tsv.", creationInfo ) 

summary(tidyFullLifeStageData$count)
hist(log(tidyFullLifeStageData$count))

WormBaseIDandGeneData <- select(tidyFullLifeStageData, c(ID, gene))
comment(WormBaseIDandGeneData) < <- paste("Mapping between WormBase IDs and 'gene' names in FASTA file")
tidyGeneIDLifeStageData <- select(tidyFullLifeStageData, -ID)
tidyWBIDLifeStageData <- select(tidyFullLifeStageData, -gene)

## Export values
## Don't use write_csv because that only works with dataframes,not tibbles
## write.csv converts the tibble to a dataframe
## For now, just save as an uncompressed .Rdata
save(tidyFullLifeStageData, mainLifeStages, tidyGeneIDLifeStageData, tidyWBIDLifeStageData, file = "processed.E-MTAB.data.Rdata", ascii=TRUE)
```

## ROC 

### Get gene names


#### Extract from FASTA file
```{r}
## Load original FASTA file
## Using Biostrings function which is not a standard df
seqData <- readDNAStringSet("c_elegan.fasta")

## names are really long descriptions.
## NEed to extract relevant part
seqDesc  <- names(seqData)


seqGeneNames <- sub(".*\\[gene=([^]]+).*", "\\1", seqDesc)

fastaInfo <- tibble(info=names(seqData), gene=seqGeneNames, length=as.double(width(seqData)))

## use gene for descript
names(seqData) <- seqGeneNames

## verify there's a match for each entry
if(sum(is.na(seqGeneNames)) ==0) print("Every entry matches")

## Verify that all 'names' are unique.
if(length(seqGeneNames) != length(unique(seqGeneNames))) print("Some genes appear twice due to isoforms")

```




#### Import Phi Values from ROC Output
```{r}
## detailed information on phi: posterior mean, posterior mean of log10(phi), etc
## StdError really StdDev of posterior
phiPosteriorInfo <- readr::read_csv("phi.posterior.unlabeled.csv") %>% dplyr::rename(phi = PHI, logPhi = log10.PHI, sd = Std.Error, logSd = log10.Std.Error)

phiData  <- bind_cols(gene=seqGeneNames, phiPosteriorInfo) #%>%
#    mutate_if(is.character, as.factor)  # convert char columns to factors

write.csv(phiData, file="labeled.phi.data.csv")
dim(phiData)
plot(phiData$phi, phiData$sd)  ## There are a bunch of anomolous genes that have a phi of 1 but an elevated sd relative to the rest of genes with that value

anomolousGenes <- phiData %>% filter(phi>0.975 & phi < 1.025 & sd>4E-5)

write.csv(anomolousGenes, "./anomolousGenes.csv")

## Plot phiData with anomolousGenes in magenta

## verify you've filtered correctly
filteredPhiData <- phiData %>% filter( !(gene %in% anomolousGenes$gene))
plot(filteredPhiData$phi, filteredPhiData$sd)



## Look at structure of data


## There are 4 entries in the "phi.mean.by.names.csv" with the name "wwp-1" 
## This is because the FASTA file contains 4 isoforms of this gene
## Problem: there's only one WB Gene ID and, thus, only one measurement.
## Solution: Combine separate estimates using median values of phi

combinedPhiData  <- phiData %>% group_by(gene) %>% summarize(median_phi = median(phi), sd_phi = median(sd), isoforms = length(phi) )
dim(combinedPhiData)
plot(combinedPhiData$median_phi, combinedPhiData$sd_phi)

filteredCombinedPhiData <- combinedPhiData %>% filter( !(gene %in% anomolousGenes$gene))

plot(filteredCombinedPhiData$median_phi, filteredCombinedPhiData$sd_phi)

save(phiData, combinedPhiData, filteredCombinedPhiData, seqData, file = "processed.ROC.data.Rdata", ascii=TRUE) 
```


#### Import Phi Values Using Lu's File -- BROKEN! NOT USED
- Don't use Lu's file
```{r eval=FALSE}

## NOTE: rocNames is corrupt.  It has replaced some gene names with date formats (e.g. apr-1 has been converted to 1-Apr
rocNamesBroken <- read_csv("phi.mean.by.names.csv", col_names = c("gene","phi2") )
rocNamesBroken[rocNamesBroken$gene=='1-Apr',]
## THere are 17 corrupted names
length(seqGeneNames) - sum(rocNamesBroken==seqGeneNames)

## Need to run 'Get estimates from ROC' code below to evaluate following line
phiBrokenDataCheck  <- bind_cols(rocNamesBroken, phiPosteriorInfo)
## Verify that phi values line up between the two datasets
plot(phiBrokenDataCheck$phi2, phiDataBrokenCheck$phi)
```


### Examine Anomolous Genes (Those with phi~1 and elevated SDs)
```{r}
## 950 genes, only 15 of these are isoform variants
length(anomolousGenes$gene)
length(unique(anomolousGenes$gene))
anomolousInfo <- filter(fastaInfo, gene %in% anomolousGenes$gene)
regularInfo <- filter(fastaInfo, !(gene %in% anomolousGenes$gene))


## Plot 2 histograms of length
## Code cribbed from: https://www.dataanalytics.org.uk/plot-two-overlapping-histograms-on-one-chart-in-r/
c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

xmax <- 800
A <- transform(anomolousInfo, n = pmin(length/3, 800) )
#anomolousInfo %>% mutate(n = min(length, 3*800))# %>% mutate(n=n/3) %>% select(n)
A <- base::as.vector(A)
B <- base::as.vector(regularInfo  %>% mutate(n = min(length, 3*800)) %>% mutate(n=n/3) %>% select(n) )


b <- min(c(A,B)) - 0.001 # Set the minimum for the breakpoints
e <- max(c(A,B)) # Set the maximum for the breakpoints
ax <- pretty(0:800, n = 16)# Make a neat vector for the breakpoints

hgA <- hist(A, breaks = ax, plot = FALSE) # Save first histogram data
hgB <- hist(B, breaks = ax,  plot = FALSE) # Save 2nd histogram data

plot(hgA, col = c1) # Plot 1st histogram using a transparent color
plot(hgB, col = c2, add = TRUE) # Add 2nd histogram using different color


```

# Modeling error in RNA Seq Data
If the underlying distribution is 
	- poisson, then Var=Rate so W = 1/count or 1/(count + k)

## Combine empirical and ROC data
```{r}

## Rename phi columns and add 'stage'
tidyFilteredAndWeightedPhiData <- filteredCombinedPhiData  %>%
    mutate(count=median_phi, weight=1/(sd_phi/mean(sd_phi))^2, stage="phi") %>%
    select(-c(isoforms, sd_phi, median_phi))


## Calculate weights for empirical data

## Combine phi
tidyFilteredAndWeightedLifeStageData <- tidyLifeStageData %>%
    filter(gene %in% standardizedPhiData$gene) %>%
    mutate(weight=(1/(count+1)) ) ## x+1 poisson weighting


tidyJoinedData <- bind_rows(tidyFilteredAndWeightedLifeStageData, tidyFilteredAndWeightedPhiData) %>%
    mutate_if(is.character, as.factor)

joinedData <- pivot_wider(tidyJoinedData, names_from=stage, values_from=c(count, weight)) 

```
## Fit Using *eivtools* 
```{r}
library(eivtools)

cData  <-bind_cols(phiData, 
eivreg(

```
